Are environmental and geographic effective surrogates for genetic variation in conservation planning?

### 2016-01-16 article/supporting_information.Rmd
use adegenet for population structure analysis instead of structure + clumpp

# download basemap
data(countriesHigh)
countries.FPLY <- countriesHigh[
	countriesHigh$ADMIN %in% c(
		'Italy', 'Switzerland', 'France', 'Austria',
		'Germany', 'Slovenia', 'Croatia', 'Hungary',
		'Monaco', 'Germany'
	)
,] %>% spFortify
```

```{r, echo=FALSE, fig.width=9.5, fig.height=10, out.width='7.5in', out.height='8.5in', fig.cap='Species distributions. Squares represent planning units. For a given species, planning units that were found to be inhabited are denoted with bright blue.'}
## plot map of species distributions
# fortify data
grid.FPLY <- spFortify(grid.PLY)
spp.grid.FPLY <- ldply(unique(spp.samples.DF$species), function(x) {
		z <- grid.FPLY[,c('long', 'lat', 'group', x),drop=FALSE]
		names(z)[4] <- 'presence'
		z$species <- gsub('\\_', ' ', x)
		return(z)
	}
)
# plot species data
ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey20', color='grey80') +
	geom_polygon(data=spp.grid.FPLY, aes(x=long, y=lat, 
		group=group, fill=presence), alpha=0.8, color='grey10') +
	theme_classic() +
	guides(fill=guide_legend(title='Presence')) +
	theme(axis.ticks=element_blank(), axis.text=element_blank(),
		axis.line=element_blank()) +
	coord_cartesian(
		xlim=buffered.range(grid.FPLY$long, 0.05),
		ylim=buffered.range(grid.FPLY$lat, 0.05)
	) +
	xlab('') +
	ylab('') +
	facet_wrap(~ species, ncol=4)
```

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6.5, fig.cap='Relative support ($\\Delta$K) for the number of populations in each species. Each panel denotes a different species. The number of populations with the highest $Delta$K value is has the most support.'}
# format data
deltak.DF <- ldply(
	seq_along(unique(spp.samples.DF$species)),
	.fun=function(i) {
		mutate(
			spp.StructureCollection.LST[[i]]@summary,
			Species=gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE)
		)
	}
)
# make plot
suppressWarnings({
	ggplot(data=deltak.DF) +
		geom_point(aes(y=delta.k, x=k)) +
		geom_line(aes(y=delta.k, x=k)) +
		theme_classic() +
		xlab('Number of populations') +
		ylab(expression(Relative~support~(Delta*K))) +
		facet_wrap(~ Species, ncol=6)
})
```

```{r, echo=FALSE, fig.width=9, fig.height=6.5, fig.cap='Population membership plots. Each panel depicts a different species. Each vertical bar represents a different individual. The colors in each bar denote the probability that the individual belongs to different populations.'}
# prepare data
popm.DF <- ldply(
	seq_along(unique(spp.samples.DF$species)),
	.fun=function(i) {
		# init
		raw.popm <- spp.StructureCollection.LST[[i]] %>% 
			structurer:::sample.membership()
		raw.popi <- apply(raw.popm, 1, which.max)
		# compile data
		curr.popm <- raw.popm %>%
			as.data.frame %>%
			mutate(id=as.character(seq_along(V2))) %>%
			gather(id, percent) %>%
			`names<-`(c('id','pop.id','percent')) %>%
			mutate(Species=gsub('\\_', ' ',
				unique(spp.samples.DF$species)[i]),
				pop.id=as.character(as.numeric(gsub('V', '', pop.id))-1))
		# sort data
		sorted.popm <- curr.popm[0,]
		for (i in seq_len(n_distinct(curr.popm$pop.id))) {
				# get individuals with greatest membership
				curr.i <- which(raw.popi==i)
				curr.DF <- filter(curr.popm,
					id %in% as.character(curr.i))
				# sort these inidividuals by membership probability
				curr.i <- curr.i[order(raw.popm[curr.i,i],
					decreasing=TRUE)]
				# rbind these individuals to sorted.popm
				for (j in curr.i)
					sorted.popm <- rbind(
						sorted.popm,
						filter(curr.DF, id==as.character(j))
					)
		}
		sorted.popm$id <- factor(as.character(sorted.popm$id), 
			levels=unique(as.character(sorted.popm$id)))
		return(sorted.popm)
	}
)
# make plot
grobLST <- dlply(popm.DF, .(Species), .fun=function(x) {
	ggplot(data=x, aes(x=id, y=percent, fill=pop.id)) +
		geom_bar(stat='identity', width=1) +
		theme_classic() +
		theme(axis.ticks=element_blank(),
			axis.text=element_blank(),
			axis.title=element_blank(),
			axis.line=element_blank(),
			legend.position='none') +
		ggtitle(x$Species[1])
})
# render plot
do.call(grid.arrange, append(grobLST, list(ncol=4,nrow=6)))
```

```{r, echo=FALSE, fig.width=6.5, fig.height=6.5, fig.cap='Distribution of populations. Each panel depicts a different species. Each pie chart represents a grid cell where individuals belonging to a species were sampled. The colors of the pie chart show the frequency of individuals belonging to different populations.'}
# prepare data
# make plot
plot(1, main='not done')
```

>>>>

# download basemap
data(countriesHigh)
countries.FPLY <- countriesHigh[
	countriesHigh$ADMIN %in% c(
		'Italy', 'Switzerland', 'France', 'Austria',
		'Germany', 'Slovenia', 'Croatia', 'Hungary',
		'Monaco', 'Germany'
	)
,] %>% spFortify
grid.FPLY <- spFortify(grid.PLY)
```

```{r, echo=FALSE, fig.width=9.5, fig.height=10, out.width='7.5in', out.height='8.5in', fig.cap='Species distributions. Squares represent planning units. For a given species, planning units that were found to be inhabited are denoted with bright blue.'}
## plot map of species distributions
# fortify data
spp.grid.FPLY <- ldply(unique(spp.samples.DF$species), function(x) {
		z <- grid.FPLY[,c('long', 'lat', 'group', x),drop=FALSE]
		names(z)[4] <- 'presence'
		z$species <- gsub('\\_', ' ', x)
		return(z)
	}
)
# plot species data
ggplot() +
	geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
		fill='grey20', color='grey80') +
	geom_polygon(data=spp.grid.FPLY, aes(x=long, y=lat, 
		group=group, fill=presence), alpha=0.8, color='grey10') +
	theme_classic() +
	guides(fill=guide_legend(title='Presence')) +
	theme(axis.ticks=element_blank(), axis.text=element_blank(),
		axis.line=element_blank()) +
	coord_cartesian(
		xlim=buffered.range(grid.FPLY$long, 0.05),
		ylim=buffered.range(grid.FPLY$lat, 0.05)
	) +
	xlab('') +
	ylab('') +
	facet_wrap(~ species, ncol=4)
```

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width=6.5, fig.height=6.5, fig.cap='Relative support ($BIC$) for the number of populations in each species. Each panel denotes a different species. Asterisks indicate the number of populations with the most support. See text for the method used to select these populations.'}
# format data
bic.buff <- laply(spp.Adegenet.LST, `[[`, 'Kstat') %>%
	c() %>% range() %>% diff %>% `*`(0.05)
bic.DF <- ldply(
	seq_along(unique(spp.samples.DF$species)),
	.fun=function(i) {
		data.frame(
			k=as.numeric(gsub('K=', '', names(spp.Adegenet.LST[[i]]$Kstat), fixed=TRUE)),
			stat=unname(spp.Adegenet.LST[[i]]$Kstat),
			best_k=as.numeric(gsub('K=', '', names(spp.Adegenet.LST[[i]]$stat), fixed=TRUE)),
			best_stat=unname(spp.Adegenet.LST[[i]]$stat) + bic.buff,
			asterisk_col=c('black',rep('transparent', length(spp.Adegenet.LST[[i]]$Kstat)-1)),
			Species=gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE)
		)
	}
)
# make plot
suppressWarnings({
	ggplot(data=bic.DF) +
		geom_point(aes(y=stat, x=k)) +
		geom_point(aes(y=best_stat, x=best_k), size=4, shape=8) +
		geom_line(aes(y=stat, x=k)) +
		theme_classic() +
		xlab('Number of populations') +
		ylab('BIC') +
		facet_wrap(~ Species, ncol=6)
})
```

```{r, echo=FALSE, fig.width=6.5, fig.height=6.5, fig.cap='Distribution of populations. Each panel depicts a different species. Squares denote planning units. Points denote the location of sampled individuals. The color of each point indicates the population to which each indiviudal was assigned.'}
# define function
plotLST <- llply(
	seq_along(unique(spp.samples.DF$species)),
	.fun=function(i) {
		# init
		curr.pts <- filter(
			spp.sample.PTS@data,
			species==unique(spp.samples.DF$species)[i]
		) %>% mutate(
			population=spp.Adegenet.LST[[i]]$grp
		)
		# create plot
		ggplot() +
			geom_polygon(data=countries.FPLY, aes(x=long, y=lat, group=group),
				fill='grey20', color='grey80') +
			geom_polygon(data=grid.FPLY, aes_string(x='long', y='lat', 
				group='group'), fill='transparent', color='grey10') +
			geom_point(data=curr.pts, aes_string(x='sample.longitude', y='sample.latitude', color='population'),
				alpha=0.3, size=3, position='jitter') +
			guides(fill=guide_legend(title=' ')) +
			theme_classic() +
			theme(axis.ticks=element_blank(), axis.text=element_blank(),
				plot.margin=unit(c(0,0,0,0),'cm'), axis.line=element_blank()) +
			coord_cartesian(
				xlim=buffered.range(grid.FPLY$long, 0.05),
				ylim=buffered.range(grid.FPLY$lat, 0.05)
			) +
			xlab('') +
			ylab('') +
			ggtitle(gsub('_', ' ', unique(spp.samples.DF$species)[i], fixed=TRUE))
	}
)
# render plot
do.call(
	grid.arrange,
	append(
		plotLST,
		list(ncol=6)
	)
)
```
